---
title: "EDA Assignment"
author: "Polina Sorokina"
date: "February 11, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r setup, include= FALSE, echo = TRUE, warning=FALSE}

library(dplyr)
library(ggplot2)
bikeVendors <- read.csv("bikeVendors.csv")

```

# Introduction and Data Description

For this reasearch, I was analyzing the Bike Vendor Data Set. The data consisted of 97 rows or observations and 35 columns or variables. I primarily used "dplyr" and "ggplot2" packages.



# Research Questions

Some of the research questions that this analysis will try to answer are: 

* Which bike models are the most and least expensive in both Mountain and Road Category?
* How many bikes of each category are in the data set under category2?
* On average, bikes with which type of frame are more expensive?




# Data Exploration

First steps in analyzing the data set was to give myself a glims of the data set using the "head" function. The function helped to identify which type of data each column belonged to. First 4 columns were Characters, fifth "Price" was an Integral, etc. 


```{r}
head(bikeVendors)
```


The data set appeared to be normal, without any missing values. Which was confirmed by using a sum formula, which added together all missing values, giving the total of `r sum(is.na(bikeVendors))`. 


Next, I performed a simple summary analysis of the "Price" column.

```{r}
summary(bikeVendors$price)
```

That formula has given me a basic statistic of the column. With lowest price being 415; highest 12,790; median 3,200; mean 3,954; 1st quartile 1,950; 3rd quartile 5,330. 

Next, I performed a deeper analysis to answer my research questions:



### Which bike models are the most expensive and the cheapest in both Mountain and Road Category?

To answer the first question, I have decided to create a graph chart of Bike Models and their Prices. 

```{r}
ggplot(bikeVendors, aes(x=model, y=price)) + geom_bar(fill="steelblue", stat = "identity") + labs(title = "Figure 1 : Price Graph of All Bikes")

```

The graph appeared to be very crowded and the model names were unreadable. 
However, this graph helped me identify that the top priced bikes are over $10,000. Therefore, I have subseted the data set to include the four most expensive bikes only, and then repeated the analysis.


```{r}
expensive_bikes <- bikeVendors %>% filter(price > 10000)

ggplot(expensive_bikes, aes(x=model, y=price)) + geom_bar(fill="steelblue", stat = "identity") + geom_text(aes(label=price)) + labs(title = "Figure 2: Price Graph of the Most Expensive Bikes")
```

As you can see on the graph, the top two most expensive bike models are: "Scalpel-Si Black Inc." and "Supersix Evo Black Inc.", both are priced at $12,790. 

I performed the same steps to identify the least expensive bikes in the market.

```{r}
cheap_bikes <- bikeVendors %>% filter(price < 600)

ggplot(cheap_bikes, aes(x=model, y=price)) + geom_bar(fill="steelblue", stat = "identity") + geom_text(aes(label=price)) + labs(title = "Figure 3: Price Graph of the Least Expensive Bikes")
```

The above graph has answered the second part of my first research question - the lease expensive bike on the market is "Catalyst 4" at $415. 

Answering this question made me curious, which categories do these bikes belong to. To answer that question, I have created a table with those three bike models and the variables I was interested in.

```{r}
bikeVendors %>% filter(model == "Supersix Evo Black Inc." | model=="Supersix Evo Hi-Mod Team" | model=="Catalyst 4") %>% subset(select=c(model,category1,category2,frame))
```

It has appeared that the most expensive bikes belong to a Road Category and are of "Elite Road" style. Their fram is Carbon. The chep bike's frame is Aluminum, it is Mountain category and is"Sport Style"





###  How many bikes of each category are in the data set under category2?

To identify that, I have created a simple table: 

```{r}
 bikeVendors %>% group_by(category2) %>% summarise(n=n())
```

Which showed that the data frame consists of 21 Elite Road Bikes; 19 Cross Country Race bikes; 16 Endurance Road bikes, etc.




### On average, bikes with which type of frame are more expensive?

To answer the last question, I subseted the data set to show a frame type and the price average only. After that, I've created another boxplot for better visualization.

```{r}

average_frame <- bikeVendors %>% group_by(frame) %>% summarise(mean = mean(price))

ggplot(average_frame, aes(frame, mean)) + geom_bar(fill="steelblue", stat = "identity") + labs(title = "Figure 4: Average Price per Frame") + ylab("average price ($)") + geom_text(aes(label=mean))

```

As you can see the Carbon bikes are significantly more expensive at the average of $5,339. While the average price of aluminum bike is only $1,979. 




#Conclusion 

The reasearch provided me with a better understanding of the data and the relationships among different variables within the data frame. I was able to identify the most and least expensive bikes along with which category those bikes belonged to. 
I was able to see the category spread of the bikes and which types of bikes are more expensive. 

The report is also posted here:

https://github.com/polinasorokina/dataanalytics/tree/master/Project